---
title: "GR5065 Homework 2"
date: "Due February 9, 2021 at 8PM New York Time"
author: "Pooja Patel"
output: 
  pdf_document: 
    latex_engine: xelatex
    number_sections: yes
urlcolor: blue
---

```{r, setup}
# call the set.seed function once here to make the knitting conditionally deterministic
set.seed(5210)
```

# Equilibrium Climate Sensitivity



## Drawing from the prior
```{r}
numerator<- rnorm(1000000, mean= 0, sd= .74 )
denom<- rnorm(1000000, mean= 0, sd= .8)

```


## Truncating the prior
The proportion that satisfies the condition of being in the interval [0, 10] is 47.195%. 
```{r}


ECS<- numerator/denom
satisfied <- ifelse((ECS < 0 | ECS > 10), 0, 1)
ECS_satisfied <- ECS[satisfied == 1]
proportion_satisfied<- (length(ECS_satisfied)/ 1000000) * 100

```

## Describing the truncated prior
```{r}
sd_ECS_satisfied<- sd(ECS_satisfied)
mean_ECS_satisfied<- mean(ECS_satisfied)
lower_tailtrue<- c(1.5, 2.0)
col1_and_2<- pnorm(lower_tailtrue, mean= mean_ECS_satisfied, sd= sd_ECS_satisfied)
higher_tailtrue<- c(4, 4.5, 6)
othercols<- pnorm(higher_tailtrue, mean= mean_ECS_satisfied, sd= sd_ECS_satisfied, lower.tail= FALSE)

col_name= c('P(X<1.5)', 'P(X<2.0)', 'P(X>4)', 'P(X>4.5)', 'P(X>6)')
row_name= c('Prior')
values= c(.501, .612, .08, .04595, .00576)
matrix_1= matrix(values, nrow= 1, ncol= 5)
rownames(matrix_1)<- row_name
colnames(matrix_1)<- col_name

print(matrix_1)

```

## PDF of a Ratio of Normals
Note that I integrated from 0 to 10, but increasing this range to negative to positive infinity sums up to 1, so this is a valid pdf, since its integral should equal 1. 
```{r}
dratio <- function(z, mu_X = -3.7, mu_Y = -1.6,
                   sigma_X = 0.2 * 3.7, sigma_Y = 0.5 * 1.6) {
  a <- sqrt(((1/(sigma_X)^2) * z^2) + (1/(sigma_Y)^2))
  b <- ((mu_X/(sigma_X)^2) * z) + (mu_Y/(sigma_Y)^2)
  c <- (mu_X^2/sigma_X^2)+ (mu_Y^2/sigma_Y^2)
  d <- exp((b^2 - c * (a^2))/(2 * (a^2)))
  pdf_dratio <- ((b * d)/a^3) * (1/(sqrt(2*pi) * sigma_X * sigma_Y)) * (pnorm(b/a) - pnorm(-b/a)) + 
    (1/(a^2 * pi * sigma_X * sigma_Y)) * exp(-c/2)
  return(pdf_dratio)
}
output1<- dratio(ECS)
trial<-integrate(dratio, 0, 10)
print(trial)

```

## Describing the truncated prior, part II
Notice the min and max values of the dratio function show that the prior is between 0 and 10. I applied this function to the untruncated ECS values and the values fall within limit. 
```{r}
min(dratio(ECS)) 
max(dratio(ECS))
#hopefully between 0-10, as it is. 

```


## The likelihood function components

```{r}
e <- c(Low_ii = .25, Low_iii = .35, Low_iv = .2, High_i = .75, High_ii = .65, High_iii = 0.6)
c <- c(Low_ii = 1.5, Low_iii = 1.5, Low_iv =  2, High_i = 4.0, High_ii = 4.5, High_iii = 4.5)
likelihood <- function(chi, e_j, c_j) {
  f = ((1 - 2*e_j) * (2*pnorm((chi - c_j)*sqrt(2)) - 1) + 1)/2
}


curve(likelihood(chi, e[1], c[1]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", ylab= 'P(e_j| chi)', lty = "solid")
curve(likelihood(chi, e[2], c[2]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", lty = "solid", add = TRUE)
curve(likelihood(chi, e[3], c[3]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", lty = "solid", add = TRUE)
curve(likelihood(chi, e[4], c[4]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", lty = "dotted", add = TRUE)
curve(likelihood(chi, e[5], c[5]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", lty = "dotted", add = TRUE)
curve(likelihood(chi, e[6], c[6]), from = 0, to = 7,
col = "black", ylim = 0:1, xname = "chi", lty = "dotted", add = TRUE)
```

## Posterior PDF
Note I have tried 2 different ways of graphing the posterior PDF. 
I also found the distribution of the denominator before finding the distribution of the posterior. 
```{r}

#Graph try1 

#numerator <- function(chi, e = e[1], c = c[1]) {
  #num <- likelihood(chi, e[1], c[1]) 
  #return(num)
#}

#posterior <- function(chi, e = e[1], c = c[1]) {
  #post <- numerator(chi) / (dnorm(e[1]))
  #return(post)
#}

#curve(posterior(chi, e[1], c[1]), from = 0, to = 10,
#col = "black", ylim = 0:1, xname = "chi", ylab= 'Posterior', lty = "solid")

#graph is not working. denominator? numerator??? 

```
```{r}
#Denominator Distribution
sd_e<- sd(e)
mean_e<- mean(e)
col1_and_2_e<- pnorm(lower_tailtrue, mean= mean_e, sd= sd_e)
othercols_e<- pnorm(higher_tailtrue, mean= mean_e, sd= sd_e, lower.tail= FALSE)

col_name_2= c('P(X<1.5)', 'P(X<2.0)', 'P(X>4)', 'P(X>4.5)', 'P(X>6)')
row_name_2= c('Denominator')
values_2= c(1, 1, 8.66e-54, 1.91e-69, 9.54e-129)
matrix_2= matrix(values_2, nrow= 1, ncol= 5)
rownames(matrix_2)<- row_name_2
colnames(matrix_2)<- col_name_2

print(matrix_2)
```

```{r}
#Graph Try 2: 
#In the paper, to reproduce the row asked for, the authors look at the ECS values <1.5 
#according to our data, the probability of that occuring regardless of any other conditions, is 
#the number of times of the ECS<1.5 / total ECS count. 
satisfied2 <- ifelse((ECS < 1.5 | ECS > 1.5), 1, 0)
ECS_satisfied2 <- ECS[satisfied == 1]
proportion_satisfied<- (length(ECS_satisfied2)/ 471950)
#All of them are less than 1.5 
#This means that the prior probability in the numerator is 1. 
#Thus this means posterior= likelihood * 1/ denominator calculated above. 
out<-function(chi, e_j, c_j){
  w<-likelihood(chi=c(1:10), e_j=e, c_j=c)/ values_2
  return(w)
}

gg<- out(chi, e, c)
ff<-c(gg)
plot(ff)


```
```{r}
b<- values/values_2

col_name_3= c('P(X<1.5)', 'P(X<2.0)', 'P(X>4)', 'P(X>4.5)', 'P(X>6)')
row_name_3= c('Posterior')
values_3= b
matrix_3= matrix(values_3, nrow= 1, ncol= 5)
rownames(matrix_3)<- row_name_3
colnames(matrix_3)<- col_name_3
print(matrix_3)
```

# Darts

```{r}
darts::drawBoard(new = TRUE)
```

## Drawing from a bivariate normal distribution


```{r}
tib_x<- rnorm(100, mean= 0, sd= 42.67)
price_x<- rnorm(100, mean=0, sd= 17.90)

mean_y_tib<- 0+(-.16)*(68.67/42.67)*(mean(tib_x))
var_y_tib<- (68.67^2)*(1-(-.16)^2)
mean_y_price<- 0+(-.22)*(39.13/17.90)*(mean(price_x))
var_y_price<- ((39.13)^2)*(1-(-.22)^2)

tib_y<- rnorm(100, mean=mean_y_tib, sd= (var_y_tib)^.5)
price_y<- rnorm(100, mean= mean_y_price, sd= (var_y_price)^.5)

darts::drawBoard(new = TRUE)
points(tib_x, tib_y, col='red')
points(price_x, price_y, col= 'green')




```


## Normal Prior Distributions
simpleEm-- variance of X variance of Y , and I am assuming a distribution of mean 50 and standard deviation 10 to draw from. This is because I'm a great dart player, so my standard deviation is very low, and from the paper the result of being a great dart player means that you can aim either in the center or slightly above center, hence the mean of 50. If I were believing things described by univariate normal distributions, I would use the default variance of 100,so standard deviation of 10, and mean 0. 
s.init= initial guess for marginal vairance= default of 100, which is the simplified Gaussian prediction for this model. 

```{r}
library(darts)
my_draw_x= rnorm(100, mean= 0, sd=100)
my_draw_y= rnorm(100, mean=0, sd=100)
my_draw_p= runif(100, min= -1, max= 1)
my_variance_y<- simpleEM(my_draw_y)
my_variance_x<- simpleEM(my_draw_x)
my_variance_p<- simpleEM(my_draw_p)
my_expscore_x<- simpleExpScores(my_variance_x$s)
my_expscore_y<- simpleExpScores(my_variance_y$s)
my_expscore_p<- simpleExpScores(my_variance_p$s)

x_var<- sample(my_expscore_x, 1)
y_var<- sample(my_expscore_y, 1)
ro_var<- sample(my_draw_p, 1)

x_throw<- rnorm(100, mean=0, sd= ((x_var)^.5))


meann<- 0+ ro_var* y_var/x_var*(x_throw)
sdd<- y_var^2*(1-ro_var^2)

y_throw<- rnorm(100, mean= meann, sd=(sdd))

darts::drawBoard(new = TRUE)
points(tib_x, tib_y, col='red')
points(price_x, price_y, col= 'green')
points(x_throw,y_throw, col= 'black')
```

# Scoring Function

```{r}
score <- function(x, y) {
  stopifnot(is.numeric(x), length(x) == 1, is.numeric(y), length(y) == 1)
  
  # convert x and y in Cartesian coordinates to a radius and angle in polar coordinates
  # https://en.wikipedia.org/wiki/Polar_coordinate_system
  radius <- sqrt(x ^ 2 + y ^ 2)
  angle  <- atan2(y, x)
  
  if (radius > 170)   return(0)  # misses dartboard
  if (radius <= 6.35) return(50) # double bullseye
  if (radius <= 15.9) return(25) # single bullseye
  
  margin <- pi / 20
  interval <- margin * 2
  small <- pi / 2 - margin - 0:19 * interval
  large <- pi / 2 + margin - 0:19 * interval
  bed <- which(angle > small & angle <= large)
  if (length(bed) == 0) {
    angle <- angle - 2 * pi
    bed <- which(angle > small & angle <= large)
  }
  S <- darts:::getConstants()$S  # 20, 1, ..., 5
  score <- S[bed]
  if (dplyr::between(radius, 99, 107))       score <- 3 * score # in triple ring
  else if (dplyr::between(radius, 162, 170)) score <- 2 * score # in double ring
  return(score)
}
z<-mapply(score, x_throw, y_throw)
hist(z)
```

## Estimating the parameters
We can see that the first values are similar, the second are very different (the variance of y expected is much higher than my actual variance), and the last value is lower for the generalEM model than my random distribution. Overall though the x variance is also a similar amount of higher than my own. I predict that the more draws I do, such as 1,000,000 for my own throws, the better and more accurate ths generalEM function will match my own random draw. Note that the correlations between the draws are stronger for my own draw than that of the generalEM draw. 

```{r}
param<- generalEM(z)
print(param$Sig.final)

```
```{r}
listy<- c(x_var^2, y_var^2, ro_var*x_var*y_var)
print(listy)
```

## Expected scores
You can see that these dots are basically on top of each other, making me aim as a regular -okay- dart player should aim, above the center as told by the paper. 
```{r}

darts::drawBoard(new = TRUE)
h<-generalExpScores(param$Sig.final)
g<-drawAimSpot(h, col= 'blue')
h_prime<-generalExpScores(listy)
g_prime<-drawAimSpot(h_prime, col= 'green')
```


