---
title: "HW1_GR5065"
author: "Pooja Patel"
date: "1/25/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Question 1
The xth fib. number is returned by this function: 
```{r}
set.seed(7)
F <- function(x) {
  stopifnot(is.numeric(x), all(x == as.integer(x)))
  sqrt_5 <- sqrt(5)
  golden_ratio <- 0.5 * (1 + sqrt_5)
  return(round(golden_ratio ^ (x + 1) / sqrt_5))
}
```


### 1.1 
The probability of knocking x/10 pins: 
```{r}
Pr<-function(x, n = 10) return(ifelse(x > n, 0, F(x)^2) / (F(n)*F(n + 1)))
  

```

### 1.2 
I know that it will return a probability of zero if x is greater than n, which means if the amount of pins capable of being knocked down, x, exceeds the amount of pins available, n, then that probability is logically 0. This is because of the ifelse statement embedded into the function. Furthermore, below, you will see a first roll and I will include the total so that you can see all the probabilities must add up to 1. Naturally, having high probabilities for knocking all 10 pins down, for example, may show a very 'good' bowler, depending on the specification of the function Pr. Also, this function is mildly realistic because the pins are aligned in a triangular fashion, much like the fibinacci sequence is when written in its traingular form; the sequence is also the sum of the previous two numbers after the first number is 0 and the second is 1, so the sequence is a simple model of how bowling works. 

### 1.3.1
```{r}
Omega <- 0:10
names(Omega) <- as.character(Omega)
round(c(Pr(Omega), total = sum(Pr(Omega))), digits = 4)  
```

The probability above must add up to 1, meaning that one of these outcomes in omega is guaranteed to happen. 

Below is one realization of our random variable for bowling, the first roll of a frame of bowling: 
```{r}

x_1 <- sample(Omega, size = 1, prob = Pr(Omega)) 
print(x_1)
```


### 1.3.2
The new sample space is 0 up until the number of pins that are remaining.- conditional probability
```{r}
Omega_2<- 0:(10-x_1)
x_2<- sample(Omega_2, size=1, prob= Pr(Omega_2, 10-x_1))
print(x_2)
```



### 1.3.3
P(x_1 pins AND x_2 pins) below-- joint probability, from the output is 8.24165 * e^ -.03 (the intersection of x=6 and then x=4 pins for roll 1 and then roll 2)
```{r}
joint_Pr <- matrix(0, nrow = length(Omega), ncol = length(Omega))
rownames(joint_Pr) <- colnames(joint_Pr) <- names(Omega)
for (x_1 in Omega) {
  Pr_x_1 <- Pr(x_1, n = 10)
  for (x_2 in Omega_2)
    joint_Pr[x_1 + 1, x_2 + 1] <- Pr_x_1 * Pr(x_2, n = 10 - x_1)
}
print(joint_Pr)
```


### 1.3.4 
I have commented below the code I tried to work on for this problem set up. 
I understood the if/else structure, but wasn't sure how to save everything into one neat place. 
```{r}
#first_rolls<- matrix(0, nrow= length(Omega), ncol= length(Omega))
#rownames(first_rolls)<- colnames(first_rolls)<- names(Omega)
# put all the first frames from 1-9 into this list called first_rolls 
#output_first_rolls<- replicate(first_rolls, sample(Omega, size= 1, prob= Pr(Omega)))
#if you knock all the pins down on your first roll for the tenth frame,
#roll again with all the pins set up. if that is also 10, then roll again (a)
# if that second roll is NOT fully 10, then a is you rolling with whatever pins where left up
#if(output_first_rolls[10]== 10){
#extra_roll1= sample(Omega, size= 1, prob= Pr(Omega)) 
  #if(extra_roll1==10){
  #a<- sample(Omega, size= 1, prob= Pr(Omega))
  #print(a)
  #}else{
  #a<- sample(0:(10-extra_roll1), size= 1, prob= Pr(0:(10-extra_roll1), 10-extra_roll1)
  #print(a)
  #}
  #else if (#the first roll plus second roll is 10 for the last frame) {
  #extra_roll1= sample(Omega, size= 1, prob= Pr(Omega))
  #print(extra_roll1)

#}
#}
```



### 1.3.5

If the above code was complete, then the answer would be the joint probabilities, which meant multiplying the probability of each amount of pins being knocked down to each other in whichever sequence the above code conducted the turn. 

### Question 2 

### 2.1 
No, I do not think it was rational for him to intially call Marle Corderio's bet. 
Notice, if the pot odds are equal to the odds of losing, the expected change in money is zero. Thus if the pot odds are greater than the odds of losing, the expected change in money should be positive, and if the pot odds are less than the odds of losing, the expected change in money should be negative. 
Suppose that for Postle, the expected change was 0. 
If this was the case, then E= 125/360 * 235 + 235/360 + -125 = 0. 
But from simple computation, we see that E in Postle's case here is -42.75, which is less than zero, significantly so. 
So technically, if we are working with optimal game theory strategy, based on this heuristic, Postle should have folded, since he was expecting to lose money, assuming, of course, that he did not know his opponent's cards.

### 2.2 

Pr(straight | a queen and a 10)= 
First, what is the raw probability of getting a straight? 
There are 10 possible sequences of straights out of a 52 card deck: 1 2 3 4 5, 2 3 4 5, etc. 
Thus having a raw straight, is 
10 * 4 * 4* 4 * 4 * 4
Subtracting out the straight flushes and raw flushes, is 10* 4 = 40 
So the total amount of combinations is 10,200 hands. 
The total amount of hands in a game of poker: 52!/ (5! * 47!)
The first number divided by the total amount of hands in a game of poker: 5/ 1274 = .39% 

P(straight | a queen and a 10)= (P(straight) * P(a queen and a 10 | a straight))/ P(a queen and a 10)

Now, P(a queen and a 10) = 4/52 * 4/51 
P(a queen and a 10 | straight)= 1

So P(straight| a queen and a 10)= (5/1274 )/ (16/51 * 4/51)= 15.95% 


### 2.3 
Bayesian statistics are mainly reliant on the fact that a conditional probability is conditional on some sort of prior information. The more information that a bayesian has in a situation, the more accurate a  person's probability calculation, thus a more accurate expectation can be calculated to facilitate in game theory decision making. If a first player checks, the other player may raise and then the other player can also check or raise, which means the first player may not lose out since he/she still has the chance to back away. But, if a fister player raises, and the second player raises even more, this means either the second player is bluffing or based on basic game theory, has a higher hand, and the first player cannot take away the amount of money he/she first raised. Thus without knowing how much the other player is willing to pay at this round of the game, the first player misses out on key information. Whereas the second player gets to see which moves the first player makes, recalibrate the probabilities and expectations, and then make a move. 

### 2.4 
The four possible combinations of hands that Cordeiro could have as hole cards, irrespective of betting, are a 4 and a 4, a 4 and a Jack, a 4 and a nine, or a 4 and an 8. 
The 4 with a 4 has a different conditional probability than the other three cases. 
 P(4|4)= P(4) * P(4|4) / P(4) = 3/51 
 All others have the same pattern as any one of them, for example:
 P(4|J) = P(4) * P(J|4) / P(J) = (4/52 * 4/51) / 4/52 = 4/51. 
 So any of those 4 events could happen, so using the "or" rule of probability: 
 3/51 + 3 * 4/51= 15/51 = 29.41% 

Considering that Cordeiro bet $200, and there is a game theory optimal consensus to raise with more than half the pot size (200 is about 55% of the pot size), then it is likely that Cordeiro has them, if she is being game theory optimal. It is more likely that she has a pair of fours, which is the more powerful hand, and relatively less likely that she has a 4 with a 9, 8, or Jack compared to the pair of fours, only because she bet that much of the pot size, which is usually done when one has a straight rather than a two pair or three pair. 

### 2.5 
The fact that Michael Postle folded does not make sense because his odds of winning were significantly higher on this turn, considering that the cards were a 9, 8, J, and 4, and Michael had a jack in his hand. In this case, he already had one pair of a jack, and getting a 10 or a jack on the next would push him into respectively a win or a tie. The probability of that happen given the cards that were dealt grows higher as more cards are dealt. 
Cordeiro's probability of winning given the cards that are dealt is naturally much higher than Michael Postle's, but when comparing it to the pot, it was more game theory optimal for Postle to not fold. His odds of winning were much higher if he bet over 50% of the pot currently than if he folded; game theory optimal means every movement towards the optimal leaves one's expected earnings higher. Any movement more towards that 50% of the pot for a bet would have increased Postle's expected return. Thus even 20% of the pot or 60% of the pot would have been better than betting 0% by folding. 
Rationally, not using this strategy is suboptimal, which is why the fact that Postle still won the game after employing such a suboptimal strategy, and does so on a consistent basis, is what made Veronica say it did not make sense. 

### 2.6 
This inference is definitely more Frequentist. For one, there is a clear independent and dependent variable: the VPIP and the amount of winnings. Secondly, one can plot some sort of line through the blue and claim that the red is an outlier, and use that argument as one to claim that Postle is cheating or not cheating. The inference from this plot that Michael Postle is cheating solely comes from the fact that Postle's scores vary largly from the trend, and a standard normal distribution of most of these points would clearly show Postle in the outer tails of the distribution. Frequentists also tend to link correlation and try to argue for causality; given a regression analysis of this data, a frequentist would try to contorl for omitted variables and minimize variance and claim that the general coefficient of winning based on the VPIP percentage is much lower than Postle's coefficient, showing that Postle is a deviation from the norm, and using this as support for the cheating hypothesis. This data is not about the cards that were given and the moves that were made given the cards; this data also does not weigh in probabilities as a bayesian would. It would be a more bayesian graph if one of the axes was the probability of winning given a low VPIP percentage versus the probability of winning given a high VPIP percentage. 

### 2.7 

The prior probability of cheating is low. 
But now that we see that Michael may be able to see his phone, and may have access to the RFID reader, then Michael's probability is now 
P(using phone to cheat | he is cheating)= 
P(he's using his phone to cheat) * P(cheating | using his phone to cheat) / P(he is cheating). 
The probability that he is using his phone to cheat increases now that we know he was using his phone. The probability that he is cheating, given that he is using his phone to cheat, is 1. 
The denominator is the overall probability (unconditional) that he is cheating. 
The main term increase is from the probability that he is using his phone to cheat, since now we have evidence that he is using his phone. 
The probability of cheating overall, as stated in the problem, is low and should remain low if all things other than the usage of the phone have stayed equal. 


### 2.8 
Bayes' Rule helps determine the probability of an event occuring, given that other things were present. This can help determine the probability that there was cheating, given the evidence. It can also help determine the probability of defmaation/libel/slander, given other evidence. Thus Bayes' rule is powerful in showing how likely an event happens; however, it still is a probability. Unless the probability is 100% or very close to 0%, one would not want to wrongly indict an innocent or acquit a guilty person, so Bayes' rule cannot be the only reason someone is indicted or made guilty. 
However, it is agreed upon in most jury situations that having as close as possible to perfect information can aid a jury come to the correct verdict. If Bayes' Rule helps bridge that gap by showing the jury the liklihoods of the occurance of the crime, then it is powerful. However, if Bayes' rule states liklihoods but people are swayed by higher probabilities and lower probabilities, then it could lead to wrongful indictment. A jury may, for example, think Postle was cheating with 95% probability, but on the 5% chance he wasn't, he could be sued and lose money for being innocent. 

Bayes' rule could be very powerful in finding probabilities to calculate expectations in terms of civil lawsuits, because usually the jury needs to come up with a number for which the sued has to pay up if he is guilty. Weighting the expected value needed to pay by the probability the person committed the crime may turn out more fair numbers monetarily speaking in the lawsuits at hand. 


